Windowed least-squares model reduction for dynamical systems

### 2019-10-14 section3/space_time.tex
minor fixes in response to comments

%Inserting the approximation~\eqref{eq:stapprox} into the optimization problem~\eqref{eq:tclsrm}, the
%constrained minimization problem over the $n$th time slab can be re-written as,
%\begin{equation}\label{eq:obj_gen_slab_spacetime} \stgenstate^n =
%\underset{\stgenstatey}{\text{argmin }} \mathcal{J}^n \bigg( \stbasisArgnt{n} \stgenstatey + \stateInterceptArg{n}  \bigg) , \end{equation}
%%\begin{equation}\label{eq:obj_gen_slab} \genstate^n(t) =
%%\underset{\genstatey}{\text{argmin }} \mathcal{J}^n
%%\bigg(\decoder\big(\genstatey(\tau)\big) \bigg) , \end{equation}
%subject to the boundary conditions 
%\begin{equation}\label{eq:bcs_spacetime}
%\stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
%\mathbb{P}^n\stbasisArg{n-1}{\timeEndArg{n-1}}\stgenstateArg{n-1}  & n = 2,\ldots,\nslabs, \\ 
%\mathbb{P}^n \stateFOMIC
% & n=1. \end{cases} \end{equation}
\subsubsection{Stationary conditions} 
The key difference between \spaceTimeAcronym\ and \spatialAcronym\ trial
	subspaces is as follows: generalized coordinates for \spaceTimeAcronym\ trial subspaces 
	comprise a vector in $\RR{\stdimArg{n}}$, while generalized coordinates for
	the \spatialAcronym\ trial subspaces comprise a
	time-dependent vector in
	$\RR{\romdimArg{n}} \otimes \timeSpaceArg{n}$. 
Thus, in the \spaceTimeAcronym\ case, the optimization problem is no longer
	minimizing a functional with respect to a (time-dependent) function, but now is minimizing a function with 
respect to vector.  As such, the first-order optimality conditions can be
	derived using standard calculus. Differentiating the objective with respect
	to the generalized coordinates and setting the result equal to zero yields 
\begin{equation}\label{eq:st_stationary}
 \intSlabArg{n} \bigg[ \stbasisDotArg{n}{t}^T  - \stbasisArg{n}{t}^T \bigg[\frac{\partial
\velocity}{\partial \stateyDiscrete} (\stbasisDotArg{n}{t} \stgenstateArg{n} +                    
\stateInterceptSTArg{n},t)\bigg]^T  \bigg] \stweightingMatArg{n} \bigg( \stbasisDotArg{n}{t} \stgenstateArg{n}  - \velocity (\stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptSTArg{n},t) \bigg) dt = \bz.\end{equation}
%subject to the boundary conditions 
%\begin{equation}\label{eq:bcs_spacetime}
%\stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
%\mathbb{P}^n\stbasisArg{n-1}{\timeEndArg{n-1}}\stgenstateArg{n-1}  & n = 2,\ldots,\nslabs, \\ 
%\mathbb{P}^n \stateFOMIC &
%n=1. \end{cases} \end{equation}
Thus, for \spaceTimeAcronym\ trial subspaces, the stationary conditions
	comprise an algebraic system of equations, as opposed to a system of
	differential equations as with \spatialAcronym\ trial subspaces.

>>>>

%Inserting the approximation~\eqref{eq:stapprox} into the optimization problem~\eqref{eq:tclsrm}, the
%constrained minimization problem over the $n$th time slab can be re-written as,
%\begin{equation}\label{eq:obj_gen_slab_spacetime} \stgenstate^n =
%\underset{\stgenstatey}{\text{argmin }} \mathcal{J}^n \bigg( \stbasisArgnt{n} \stgenstatey + \stateInterceptArg{n}  \bigg) , \end{equation}
%%\begin{equation}\label{eq:obj_gen_slab} \genstate^n(t) =
%%\underset{\genstatey}{\text{argmin }} \mathcal{J}^n
%%\bigg(\decoder\big(\genstatey(\tau)\big) \bigg) , \end{equation}
%subject to the boundary conditions 
%\begin{equation}\label{eq:bcs_spacetime}
%\stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
%\mathbb{P}^n\stbasisArg{n-1}{\timeEndArg{n-1}}\stgenstateArg{n-1}  & n = 2,\ldots,\nslabs, \\ 
%\mathbb{P}^n \stateFOMIC
% & n=1. \end{cases} \end{equation}
\subsubsection{Stationary conditions} 
The key difference between \spaceTimeAcronym\ and \spatialAcronym\ trial
	subspaces is as follows: generalized coordinates for \spaceTimeAcronym\ trial subspaces 
	comprise a vector in $\RR{\stdimArg{n}}$, while generalized coordinates for
	the \spatialAcronym\ trial subspaces comprise a
	time-dependent vector in
	$\RR{\romdimArg{n}} \otimes \timeSpaceArg{n}$. 
Thus, in the \spaceTimeAcronym\ case, the optimization problem is no longer
	minimizing a functional with respect to a (time-dependent) function, but now is minimizing a function with 
respect to vector.  As such, the first-order optimality conditions can be
	derived using standard calculus. Differentiating the objective with respect
	to the generalized coordinates and setting the result equal to zero yields 
\begin{equation}\label{eq:st_stationary}
 \intSlabArg{n} \bigg[ \stbasisDotArg{n}{t}^T  - \stbasisArg{n}{t}^T \bigg[\frac{\partial
\velocity}{\partial \stateyDiscrete} (\stbasisDotArg{n}{t} \stgenstateArg{n} +                    
\stateInterceptSTArg{n},t)\bigg]^T  \bigg] \stweightingMatArg{n} \bigg( \stbasisDotArg{n}{t} \stgenstateArg{n}  - \velocity (\stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptSTArg{n},t) \bigg) dt = \bz.\end{equation}
%subject to the boundary conditions 
%\begin{equation}\label{eq:bcs_spacetime}
%\stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
%\mathbb{P}^n\stbasisArg{n-1}{\timeEndArg{n-1}}\stgenstateArg{n-1}  & n = 2,\ldots,\nslabs, \\ 
%\mathbb{P}^n \stateFOMIC &
%n=1. \end{cases} \end{equation}
Thus, for \spaceTimeAcronym\ trial subspaces, the stationary conditions
	comprise a system of algebraic equations, as opposed to a system of
	differential equations as with \spatialAcronym\ trial subspaces.

### 2019-10-14 section3/tclsrm_formulation.tex
minor fixes in response to comments

\subsection{\methodAcronym\ summary}
This section outlined the \methodAcronym\ method for model reduction. In
summary, \methodAcronym\ sequentially minimizes the time continuous FOM ODE
residual within the range of a space--time trial over time
windows subject to boundary conditions associated with continuity across time
windoes; i.e.,  
%\methodAcronym\  
%computes the generalized coordinates
%$\genstateArgnt{n}$, for $n = 1,\ldots,\nslabs$, as the solution to the
%minimization problems, 
\methodAcronym\ sequentially computes approximate solutions
$\approxstateArgnt{n}\in\stspace^n$, $n=1,\ldots,\nslabs$ that comprise
solutions to the optimization problem \eqref{eq:tclsrm}.
\KTC{I removed repeated equations here and just referred to their previous
definitions. This will cut down on the text.}
%\begin{equation*}
%\underset{\stateyArgnt{} \in \stspace^n}{\text{minimize } }
%\mathcal{J}^n(\stateyArgnt{}),
%\end{equation*}
%where
%\begin{align*}
%\mathcal{J}^n &\vcentcolon \statey \mapsto
%\frac{1}{2} \int_{\timeStartArg{n}}^{\timeEndArg{n}} \big[ \dot{\statey}(t)
%- \velocity(\stateyArg{}{t},t) \big]^T \stweightingMatArg{n} \big[
%\dot{\statey}(t) - \velocity(\stateyArg{}{t},t) \big] d t \\
%&\vcentcolon \RR{\fomdim}\otimes \timeSpaceArg{n} \rightarrow
%\RRplus,
%\end{align*}
%subject to the boundary conditions.

For \spatialAcronym\ trial subspaces, the stationary conditions for the
residual minimization problem can be derived via the Euler--Lagrange equations
and yield the system \eqref{eq:lspg_continuous}--\eqref{eq:lspg_bcs} over the $n$th window for $t \in
[\timeStartArg{n},\timeEndArg{n}]$.
%\begin{align*}\
%&\massArg{n}  \genstateDotArg{n}{t}  -  \basisspaceTArg{n}
%\stweightingMatArgt{n}{t} \velocity(\veloargsromn) =  \massArg{n} \adjointArg{n}{t} , \\
% &\massArg{n} \adjointDotArg{n}{t}  + \basisspaceTArg{n} \bigg[\frac{\partial
%\velocity}{\partial \statey} (\veloargsromn) \bigg]^T \stweightingMatArgt{n}{t} \basisspaceArg{n}
% \adjointArg{n}{t} = \nonumber \\ 
%&-\basisspaceTArg{n} \big[
%\frac{\partial \velocity}{\partial \statey}(\veloargsromn) \big]^T \stweightingMatArgt{n}{t} \big( \mathbf{I} -
%\basisspaceArg{n} [\massArg{n}]^{-1} \basisspaceTArg{n} \stweightingMatArgt{n}{t} \big)
% \big( \basisspace \genstateDotArg{n}{t}  -
%\velocity(\veloargsromn) \big),  &
%%\genstateArg{n}{\timeStartArg{n}} = \begin{cases}
%%\genstate^{n-1}(\timeEndArg{n-1}) & n=2,\ldots,\nslabs, \\
%%\basisspace^T(\approxstate_0 - \stateIntercept) & n=1, \end{cases}\\ &
%%\adjointArg{n}{\timeEndArg{n}} = \boldsymbol 0.
%\end{align*} 
%subject to the boundary conditions~\eqref{eq:lspg_bcs1}--\eqref{eq:lspg_bcs}. 
\methodAcronym\ with \spatialAcronym\ trial subspaces can be alternatively interpreted as a Lagrange problem from optimal control: the Galerkin method is forced by a 
controller that enforces the residual minimization property. 
Section~\ref{sec:wls_spacetime} additionally considered \spaceTimeAcronym\
trial subspaces. For \spaceTimeAcronym\ trial subspaces, in where the
generalized coordinates are no longer functions, the stationary conditions
correspond to the system of algebraic equations \eqref{eq:st_stationary}.

>>>>

\subsection{\methodAcronym\ summary}
This section outlined the \methodAcronym\ \approachKwd\ for model reduction. In
summary, \methodAcronym\ sequentially minimizes the time continuous FOM ODE
residual within the range of a space--time trial subspace over time
windows; i.e.,  
%\methodAcronym\  
%computes the generalized coordinates
%$\genstateArgnt{n}$, for $n = 1,\ldots,\nslabs$, as the solution to the
%minimization problems, 
\methodAcronym\ sequentially computes approximate solutions
$\approxstateArgnt{n}\in\stspace^n$, $n=1,\ldots,\nslabs$ that comprise
solutions to the optimization problem \eqref{eq:tclsrm}.
%\KTC{I removed repeated equations here and just referred to their previous
%definitions. This will cut down on the text.}
%\begin{equation*}
%\underset{\stateyArgnt{} \in \stspace^n}{\text{minimize } }
%\mathcal{J}^n(\stateyArgnt{}),
%\end{equation*}
%where
%\begin{align*}
%\mathcal{J}^n &\vcentcolon \statey \mapsto
%\frac{1}{2} \int_{\timeStartArg{n}}^{\timeEndArg{n}} \big[ \dot{\statey}(t)
%- \velocity(\stateyArg{}{t},t) \big]^T \stweightingMatArg{n} \big[
%\dot{\statey}(t) - \velocity(\stateyArg{}{t},t) \big] d t \\
%&\vcentcolon \RR{\fomdim}\otimes \timeSpaceArg{n} \rightarrow
%\RRplus,
%\end{align*}
%subject to the boundary conditions. 
For \spatialAcronym\ trial subspaces, the stationary conditions for the
residual minimization problem can be derived via the Euler--Lagrange equations
and yield the system \eqref{eq:lspg_continuous}--\eqref{eq:lspg_bcs} over the $n$th window for $t \in
[\timeStartArg{n},\timeEndArg{n}]$.
%\begin{align*}\
%&\massArg{n}  \genstateDotArg{n}{t}  -  \basisspaceTArg{n}
%\stweightingMatArgt{n}{t} \velocity(\veloargsromn) =  \massArg{n} \adjointArg{n}{t} , \\
% &\massArg{n} \adjointDotArg{n}{t}  + \basisspaceTArg{n} \bigg[\frac{\partial
%\velocity}{\partial \statey} (\veloargsromn) \bigg]^T \stweightingMatArgt{n}{t} \basisspaceArg{n}
% \adjointArg{n}{t} = \nonumber \\ 
%&-\basisspaceTArg{n} \big[
%\frac{\partial \velocity}{\partial \statey}(\veloargsromn) \big]^T \stweightingMatArgt{n}{t} \big( \mathbf{I} -
%\basisspaceArg{n} [\massArg{n}]^{-1} \basisspaceTArg{n} \stweightingMatArgt{n}{t} \big)
% \big( \basisspace \genstateDotArg{n}{t}  -
%\velocity(\veloargsromn) \big),  &
%%\genstateArg{n}{\timeStartArg{n}} = \begin{cases}
%%\genstate^{n-1}(\timeEndArg{n-1}) & n=2,\ldots,\nslabs, \\
%%\basisspace^T(\approxstate_0 - \stateIntercept) & n=1, \end{cases}\\ &
%%\adjointArg{n}{\timeEndArg{n}} = \boldsymbol 0.
%\end{align*} 
%subject to the boundary conditions~\eqref{eq:lspg_bcs1}--\eqref{eq:lspg_bcs}. 
\methodAcronym\ with \spatialAcronym\ trial subspaces can be alternatively interpreted as a Lagrange problem from optimal control: the Galerkin method is forced by a 
controller that enforces the residual minimization property. 
Section~\ref{sec:wls_spacetime} additionally considered \spaceTimeAcronym\
trial subspaces. For \spaceTimeAcronym\ trial subspaces, in where the
generalized coordinates are no longer functions, the stationary conditions
correspond to the system of algebraic equations \eqref{eq:st_stationary}.

### 2019-10-14 section4/numerical_methods.tex
minor fixes in response to comments

\section{Numerical solution techniques}\label{sec:numerical_techniques}
Given the formulation of the \methodAcronym\ method presented in Section
\ref{sec:tclspg}, we now consider numerical-solution techniques for the method. 
Due to the commonalities that the \methodAcronym\ method exhibits with
optimal-control problems, \KTC{is this only for S-reduction subspaces? If so,
state this.} the derivation of these techniques can leverage a large existing
body of work. In this section, we outline a variety of such techniques. We
focus primarily on 
direct (i.e., discretize-the-optimize) and indirect (i.e., optimize-then-discretize) for \spatialAcronym\ trial subspaces, 
and then briefly outline direct and indirect methods for \spaceTimeAcronym\ trial subspaces.

\subsection{\spatialAcronym: direct and indirect methods}
Numerical techniques to
solve optimal-control problems can be classified as either
\textit{direct} or \textit{indirect}
methods~\cite{conway_optimalcontrolreview}. Rather than working with the first order optimality conditions~\eqref{eq:lspg_continuous}--\eqref{eq:lspg_bcs}, direct methods ``directly" solve the optimization problem~\eqref{eq:tclsrm} by first
numerically discretizing the system and objective functional and ``transcribing"
the infinite dimensional problem into a finite-dimensional nonlinear
programming (optimization) problem. Direct approaches thus
``discretize-then-optimize''.
In contrast, indirect methods compute solutions to the Euler--Lagrange equations~\eqref{eq:lspg_continuous}--\eqref{eq:lspg_bcs} (which comprise the first-order optimality conditions of the objective functional). Thus, indirect methods ``optimize-then-discretize," and solve the optimization problem
``indirectly". A variety of both discretization/solution \KTC{why are these
lumped together? Could we just say ``discretization'', ``solution'', or
``discretization and solutin''?} techniques are possible
for both direct and indirect methods. Collocation methods,
finite-element methods, spectral methods, and shooting methods are all examples of
plausible solution techniques.

>>>>

\section{Numerical solution techniques}\label{sec:numerical_techniques}
We now consider numerical-solution techniques for the \methodAcronym\ \approachKwd.
We focus primarily on 
direct (i.e., discretize-the-optimize) and indirect (i.e., optimize-then-discretize) for \spatialAcronym\ trial subspaces, 
and then briefly outline direct and indirect methods for \spaceTimeAcronym\ trial subspaces. 
%It is noted that, as \methodAcronym\ with \spatialAcronym\ trial subspaces 
%Given the formulation of the \methodAcronym\ method presented in Section
%\ref{sec:tclspg}, we now consider numerical-solution techniques for the method. 
%Due to the commonalities that the \methodAcronym\ method exhibits with
%optimal-control problems, \KTC{is this only for S-reduction subspaces? If so,
%state this.} the derivation of these techniques can leverage a large existing
%body of work. In this section, we outline a variety of such techniques. 
\subsection{\spatialAcronym: direct and indirect methods}
\methodAcronym\ with \spatialAcronym\ trial subspaces can be viewed as an optimal-control problem.
Numerical techniques to
solve optimal-control problems can be classified as either
\textit{direct} or \textit{indirect}
methods~\cite{conway_optimalcontrolreview}. Rather than working with the first order optimality conditions~\eqref{eq:lspg_continuous}--\eqref{eq:lspg_bcs}, direct methods ``directly" solve the optimization problem~\eqref{eq:tclsrm} by first
numerically discretizing the system and objective functional and ``transcribing"
the infinite dimensional problem into a finite-dimensional nonlinear
programming (optimization) problem. Direct approaches thus
``discretize-then-optimize''.
In contrast, indirect methods compute solutions to the Euler--Lagrange equations~\eqref{eq:lspg_continuous}--\eqref{eq:lspg_bcs} (which comprise the first-order optimality conditions of the objective functional). Thus, indirect methods ``optimize-then-discretize," and solve the optimization problem
``indirectly". A variety of both discretization and solution techniques are possible
for both direct and indirect methods. Collocation methods,
finite-element methods, spectral methods, and shooting methods are all examples of
plausible solution techniques.
